---
title: "Penguin Flipper Lengths Anova"
author: "Sandra Fogg"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Load Packages and read in penguins data

```{r, include = FALSE}
library(tidyverse)
library(car)
library(onewaytests)

penguins_df <- read_csv("penguins.csv")
```

2. Is there a significant difference in mean flipper lengths for female Chinstrip, Gentoo, and Adelie Penguins near Palmer Archipelago?

There is one factor: species
There are three levels: Chinstrap, Gentoo, Adelie
There is one continous variable: flippper length

Perhaps we can use one-way ANOVA...if it meets assumptions
- Independent observations
- Equal Variance (Levene's Test)
- Normality (histograms/qqplots)

3. Find counts or "n" for each level

counts <- with(penguins, table(species, sex))
```{r}
counts <- with(penguins_df, table(species, sex))

View(counts)

# Indicates that if n>30 if using means, CLT suggests that it is normal distribution. 
# Perform other tests to see if mean is the best measure of central tendency
```
