---
title: "Penguin Flipper Lengths Anova"
author: "Sandra Fogg"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Load Packages and read in penguins data

```{r, include = FALSE}
library(tidyverse)
library(car)
library(onewaytests)

penguins_df <- read_csv("penguins.csv")
```

2. Is there a significant difference in mean flipper lengths for female Chinstrip, Gentoo, and Adelie Penguins near Palmer Archipelago?

There is one factor: species
There are three levels: Chinstrap, Gentoo, Adelie
There is one continous variable: flippper length

Perhaps we can use one-way ANOVA...if it meets assumptions
- Independent observations
- Equal Variance (Levene's Test)
- Normality (histograms/qqplots)

3. Find counts or "n" for each level

counts <- with(penguins, table(species, sex))
```{r}
counts <- with(penguins_df, table(species, sex))

View(counts)

# Indicates that if n>30 if using means, CLT suggests that it is normal distribution. 
# Perform other tests to see if mean is the best measure of central tendency
```

# Create a new, simplified table
```{r}
penguins_df2 <- penguins_df %>%
  mutate(
    common_name = 
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo")) %>% 
      select(-species) %>% 
      filter(sex == "FEMALE")

penguins_df
```

4. Make exploratory graphs and perform Levene's test
```{r}
p_hist <- ggplot(penguins_df2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~ common_name, scale = "free")

p_hist

p_qq <- ggplot(penguins_df2, aes(sample = flipper_length))+
  geom_qq(aes(color = common_name)) +
  facet_wrap(~ common_name, scale = "free")

p_qq

# Based on qq-plots and histograms, observations for each group appear normally distributed.  Also, n > 30 for all groups, so means will be normally distrubuted by CLT. 
  
```


5. Evaluate variances and performa a Levene's Test for equal variances

```{r}
# If the largest sample variance is < 4x greater than the smallest sample variance, then these are "close enough" to consider variances equal

variances <- penguins_df2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )

# Units are squared for variance = mm^2

# The largest variance (33.1) is not 4x greater than the smallest variance (15.2)

# Complete formal hypothesis for variances (f-test for two-samples, but need to use Levene's test because there are three)

# Levene's test is in 'car' package

levene_penguin <- leveneTest(flipper_length ~ common_name, data = penguins_df2)
levene_penguin


# H0: No difference in variances ( variances are equal)
# HA: Variances are NOT equal

# There is no significant difference in variances across species
```

6. Perform a one-way ANOVA for penguin flipper length by species (Common Name)

```{r}
penguin_aov <- aov(flipper_length ~ common_name, data = penguins_df2)

summary(penguin_aov)


# Sample means are this different because they were most likely drawn for different population
# H0: Means across all groups are equal
# HA: At least two group means differ significantly

# Reject the null.  There is a difference.  Which two?  Use post-hoc testing.


# H0:  No significant difference in pairwise means
# HA: There is a significant difference in pairwise means
penguin_pH <- TukeyHSD(penguin_aov)

penguin_pH

# Reject the null in every pairwise comparision

```

Mean flipper lengths of female penguins differed significantly for all species studied (one-way anova, F(df) = statistic, p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05)
